---
title: "Data exploration"
author: "Shane Dewees"
date: "12/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(hrbrthemes)
library(here)
library(lubridate)
```

## Loading in Variables

```{r}
veg_enviro_fire_drought <- read.csv(here("csv_data", "veg_enviro_fire_drought_4.csv"))
chaparral_analysis <- read.csv(here("csv_data", "chaparral_analysis_drought_4.csv"))
```

## Secondary Cover

```{r}
ggplot(veg_enviro_fire_drought, aes(secondary_cover_1930)) +
  geom_histogram(position = "identity", bins = 20) +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = secondary_cover_1930)) +
  geom_qq()
```


## Elevation

```{r}
ggplot(veg_enviro_fire_drought, aes(elevation)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = elevation)) +
  geom_qq()
```
normal

## Aspect

```{r}
ggplot(veg_enviro_fire_drought, aes(aspect)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = aspect)) +
  geom_qq()
```

not-nomral

## Slope

```{r}
ggplot(veg_enviro_fire_drought, aes(slope)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = slope)) +
  geom_qq()
```

normal

## Solar Radiation Summer

```{r}
ggplot(veg_enviro_fire_drought, aes(solar_radiation_summer)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = solar_radiation_summer)) +
  geom_qq()
```

normal

## Solar Radiation Equinox

```{r}
ggplot(veg_enviro_fire_drought, aes(solar_radiation_equinox)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = solar_radiation_equinox)) +
  geom_qq()
```

not normal

## Solar Radiation Winter

```{r}
ggplot(veg_enviro_fire_drought, aes(solar_radiation_winter)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = solar_radiation_winter)) +
  geom_qq()
```

Not normal

## Road Distance 

```{r}
ggplot(veg_enviro_fire_drought, aes(road_dist)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = road_dist)) +
  geom_qq()
```
 
 not-normal
 
## Precipitation

```{r}
ggplot(veg_enviro_fire_drought, aes(precipitation)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = precipitation)) +
  geom_qq()
```

not-normal

## Max Temperature

```{r}
ggplot(veg_enviro_fire_drought, aes(max_temp)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = max_temp)) +
  geom_qq()
```

normal

## Mean Temperature

```{r}
ggplot(veg_enviro_fire_drought, aes(mean_temp)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = mean_temp)) +
  geom_qq()
```

not-normal

## Minimum Temperature

```{r}
ggplot(veg_enviro_fire_drought, aes(min_temp)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = min_temp)) +
  geom_qq()
```

not-normal

## Max January VPD 

```{r}
ggplot(veg_enviro_fire_drought, aes(max_vpd_jan)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = max_vpd_jan)) +
  geom_qq()
```

not-normal

## Max August VPD 

```{r}
ggplot(veg_enviro_fire_drought, aes(max_vpd_aug)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = max_vpd_aug)) +
  geom_qq()
```

not-normal

## Clay

```{r}
ggplot(veg_enviro_fire_drought, aes(clay)) +
  geom_histogram(position = "identity", bins = 10)

ggplot(veg_enviro_fire_drought, aes(sample = clay)) +
  geom_qq()
```

## Silt

```{r}
ggplot(veg_enviro_fire_drought, aes(silt)) +
  geom_histogram(position = "identity", bins = 10)

ggplot(veg_enviro_fire_drought, aes(sample = silt)) +
  geom_qq()
```

## Available Water Supply

```{r}
ggplot(veg_enviro_fire_drought, aes(available_water_supply)) +
  geom_histogram(position = "identity", bins = 10)

ggplot(veg_enviro_fire_drought, aes(sample = available_water_supply)) +
  geom_qq()
```

## Organic matter

```{r}
ggplot(veg_enviro_fire_drought, aes(organic_matter)) +
  geom_histogram(position = "identity", bins = 10)

ggplot(veg_enviro_fire_drought, aes(sample = organic_matter)) +
  geom_qq()
```

## Minimum Anomaly

```{r}
ggplot(veg_enviro_fire_drought, aes(min_anomaly)) +
  geom_histogram(position = "identity") +
  theme_classic()

ggplot(veg_enviro_fire_drought, aes(sample = min_anomaly)) +
  geom_qq()
```

categorical


## Correlations 

```{r}
variables_cor <- cor(veg_enviro_fire_drought[11:23])
pairs(veg_enviro_fire_drought[11:23])
```


